CONEM2024-0025\\ ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-02 accepted/CONEM2024/FCPCOM/CONEM2024.tex
comparative weights

@@

    \subsection{Algorithm for forcing low consistency rate}

@@

>>>>

    \section{ALGORITHM FOR FORCING LOW CONSISTENCY RATE}

@@

    \subsection{Analysis on the Algorithm for Adjusting Inconsistencies}

@@

\subsection{Reflections on the Correct Application of the Multicriteria Classification}

In our study, we observed that one of the crucial points for correctly applying multicriteria ABC classification is correctly attributing the weights for the initial ABC classification.

When we built the input data file, we considered two types of criteria for ABC analysis, cumulative sum (for the net profit and lead time criteria) and qualitative criteria measured by weights (0 and 1, in the case of the criticality criterion, and 1, 2, 3, in the case of the other criteria). As a result, we had a bias in the multicriteria classification. The criterion with the top percentage difference (turnover) was dominant.

In the case of the criticality criterion, the choice of a binary parameter results in a distribution of the probability among the criteria that are critical. This seems to be consistent with the desired goal. However, in the case of the other criteria, the consistency is no longer so clear. For example, in the company, the product with the highest turnover (242,280.00 reais) has a turnover corresponding to 12\% of the total turnover. A product with a longer lead time (30 days) represents 0.46\% of the total lead time, that is, a percentage 0.25\% higher than products with a lead time of 20 days (0.31\%) and 0.35\% higher than products with a lead time of 7 days (0.11\%).

The question is: is the difference between the importance of a product with a 30-day lead time and a product with a 7-day lead time really only 0.35\%. This importance can only really be appreciated when put in comparison with the difference in importance of 12\% between the product with the highest turnover and the product with the lowest turnover.

Analyzing these values, we come to the conclusion that it seems coherent to adopt as a measure for the weights not gross values, but comparative values. This conclusion gave rise to the idea of assigning comparative weights.
